---
title: "Skills & Deliverables"
description: "A toolbox of methods and outputs developed across MGEM."
format:
  html:
    toc: true
editor: visual
---

\
Selected skills, tools, and deliverable from my MGEM studies. \
\
\## Overview\
\
Through the MGEM program, I developed skills in:\
\
- remote sensing and raster processing\
- GIS mapping and cartography\
- LiDAR and aerial image analysis\
- spatial analysis in R\
- basic coding and data visualization\
\
---\

## FCOR 599 Capstone Project

\
\*\*Focus:\*\* Garry oak detection using LiDAR and orthophotos \
\*\*Main tools:\*\* \`terra\`, \`lidR\`, \`sf\`, \`dplyr\`, \`ggplot2\`, \`randomForest\`\
\
\### Skills\
- crown-level feature extraction\
- classification workflow design\
- model comparison\
- map production\
- interpreting accuracy results\
\
\### Example R code: simple crown classification\

```{r}
library(dplyr)

crowns <- data.frame(  
  crown_id = 1:8,  
  height_m = c(12, 18, 9, 16, 7, 20, 11, 15),  
  ndvi = c(0.72, 0.81, 0.55, 0.78, 0.49, 0.84, 0.61, 0.76))

crowns <- crowns %>%  
  mutate(prediction = ifelse(height_m > 10 & ndvi > 0.70, "Garry oak", "Other"))

crowns
```

## GEM 520 – Remote Sensing I

**Focus:** raster processing, satellite imagery, vegetation indices\
**Main tools:** `terra`, `ggplot2`

### Skills

-   raster handling

-   band math

-   NDVI calculation

-   simple classification

-   plotting raster outputs

### Example R code: NDVI with simple raster layers

```{r}
library(terra)

red <- rast(nrows = 4, ncols = 4, vals = c(
  0.2, 0.3, 0.4, 0.2,
  0.1, 0.2, 0.3, 0.2,
  0.2, 0.4, 0.5, 0.3,
  0.1, 0.2, 0.2, 0.1
))

nir <- rast(nrows = 4, ncols = 4, vals = c(
  0.6, 0.7, 0.8, 0.6,
  0.5, 0.6, 0.7, 0.5,
  0.6, 0.8, 0.9, 0.7,
  0.4, 0.5, 0.6, 0.4
))

ndvi <- (nir - red) / (nir + red)
plot(ndvi, main = "Example NDVI")
```

## GEM 521 – Remote Sensing II

**Focus:** ALS, hyperspectral thinking, feature extraction, classification\
**Main tools:** `terra`, `ggplot2`, `dplyr`

### Skills

-   working with spectral and structural variables

-   comparing classes

-   extracting meaningful features

-   classification thinking

### Example R code: comparing spectral signatures

```{r}
library(ggplot2)

spectra <- data.frame(
  wavelength = c(450, 550, 650, 850, 1600, 2200),
  vegetation = c(0.08, 0.12, 0.05, 0.52, 0.30, 0.18),
  soil = c(0.15, 0.20, 0.25, 0.30, 0.35, 0.32)
)

spectra_long <- reshape(
  spectra,
  varying = c("vegetation", "soil"),
  v.names = "reflectance",
  timevar = "class",
  times = c("Vegetation", "Soil"),
  direction = "long"
)

ggplot(spectra_long, aes(x = wavelength, y = reflectance, color = class)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  labs(title = "Example Spectral Profiles",
       x = "Wavelength (nm)",
       y = "Reflectance")
```

## GEM 511 – Cartography and Visualization

**Focus:** map design, layout, communication, visual clarity\
**Main tools:** `ggplot2`, `sf`

### Skills

-    map readability

-   symbol choices

-   clear visual communication

-   making figures more professional

### Example R code: simple thematic map style

```{r}
library(ggplot2)

areas <- data.frame(
  neighborhood = c("North", "South", "East", "West"),
  park_cover = c(12, 28, 19, 35)
)

ggplot(areas, aes(x = neighborhood, y = park_cover)) +
  geom_col() +
  labs(title = "Example Thematic Figure",
       x = "Area",
       y = "Park Cover (%)")
```

## FCOR 511 – Diverse Knowledge Systems 

**Focus:** collaboration, sustainability, land-based perspectives

### Skills

-   integrating social and environmental ideas

-   thinking beyond technical analysis

-   considering multiple values in decision-making

## Toolbox

Below are some of the main tools I used across the MGEM program:

-   R / RStudio

-   terra

-   sf

-   ggplot2

-   dplyr

-   lidR

-   ArcGIS Pro

-   QGIS

-   Python (basic use)
